---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

First read the csv and load it to the activity variable using read.csv(). 

```{r}
activity <- read.csv("activity.csv")
str(activity)
```

Then by looking at the data types of the activity variable's columns we find that the date column is character, so we transform it to date format.

```{r}
activity$date <- as.Date(activity$date)
str(activity)
```

## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day

```{r barplot, fig.height=5}
par(mar = c(6,6,2,2))
barplot(tapply(activity$steps, activity$date, sum, na.rm = TRUE), las = 2, col = "light blue", main = "Total number of steps / Day", cex.names = 0.6)
title(xlab = "Date", line = 4)
title(ylab = "Number of Steps", line = 4)
```



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
